---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# sdrUpload

<!-- badges: start -->
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip) [![Travis-CI Build Status](https://api.travis-ci.org/thomascrines/admStructuredData.svg?branch=master)](https://travis-ci.org/thomascrines/admStucturedData)
<!-- badges: end -->

admStructuredData is an R package which allows SG analysts to upload data from Excel workbooks to the Analytical Data Management (ADM) structured data repository.

This branch is specifically for testing new functionality, allowing users to create versions of datasets and retrieve data as it was at the point of the version's creation. This branch also includes function renaming, so be aware that existing pipelines may be affected by changes in function and argument names.

Data can be uploaded from a single worksheet, all worksheets in a workbook, or all workbooks in a folder. 

To use the package you will need to have access to a database in the ADM structured data repository; if you don't have this please contact the [ADM support team](mailto:adm.support@gov.scot).

## Installation

Install opendatascot from GitHub:

1. Go to the sdrUpload [repository](https://github.com/thomascrines/admStructuredData) on GitHub
2. Click **Clone or download** then **Download ZIP**
3. Save the file locally and unzip
4. Install with install.packages():

```{r source-installation, eval = FALSE}
install.packages("C:\\DownloadDirectory\\admStructuredData-master\\admStructuredData-master", repos = NULL,
                 type="source", lib = "C:\\YourLibraryPath")
```
Please see the [SG R/RStudio Instructions for users](https://erdm.scotland.gov.uk:8443/documents/A23744528/details) before installing the package.

The package will only work when installed on a SCOTS machine.

Prerequisites
---------
To use admStructuredData you will need to have access to a database in the ADM structured data repository; if you don't have this please contact the [ADM support team](mailto:adm.support@gov.scot). Most functions will require 'database' and 'server' arguments; the values for these will be provided by the ADM team when your database is set up.

Some functions allow processed files to be moved to an archive folder. To use this feature, the archive folder has to be made in advance.

Examples
---------
### Write data to database

Write an R dataframe to a database table:

`adm_upload_dataframe(database = "DatabaseName", server = "ServerName", table = "ExampleDatabaseTable", dataframe = example_dataframe)`

Write data from an Excel worksheet to a database table:

`adm_upload_worksheet(database = "DatabaseName", server = "ServerName", file_path = "C:\\Temp\\Sdr\\ExcelFile.xlsx", worksheet = "WorksheetName")`

Write data all Excel sheets contained in an Excel workbook to database tables:

`adm_upload_workbook(database = "DatabaseName", server = "ServerName\\Instance", file_path = "C:\\Temp\\Sdr\\ExcelFile.xlsx")`

Write data from all Excel sheets contained in all Excel files contained in a source directory to database tables:

`adm_upload_folder(database = "DatabaseName", server = "ServerName\\Instance", file_path = "C:\\Temp\\Sdr\\ExcelFile.xlsx")`

### Other database functions

Delete a table from a database:

`adm_delete_table(database = "DatabaseName", server = "ServerName", table = "TableName")`

Return a list of tables in a database:

`adm_list_tables(database = "DatabaseName", server = "DatabaseServer")`

Return data from a database table:

`adm_import_table(database = "DatabaseName", server = "ServerName", table = "TableName")`

### Append / overwrite arguments

`adm_upload_dataframe`, `adm_upload_worksheet`, `adm_upload_workbook` and `adm_upload_folder` all accept `append` and `overwrite` arguments.
These are optional, but if neither is set an error will be thrown if the database table already exists.
Only one can be set to `TRUE` at a time or an error will be thrown.

Append:

`adm_upload_dataframe(database = "DatabaseName", server = "ServerName", table = "example_database_table", dataframe = example_dataframe, append = TRUE)`

Overwrite:

`adm_upload_dataframe(database = "DatabaseName", server = "ServerName", table = "example_database_table", dataframe = example_dataframe, overwrite = TRUE)`

### Archive argument

`adm_upload_workbook` and `adm_upload_folder` can also accept an `archive_directory_path` argument, to move processed files from the original location to a specified directory:

`adm_upload_folder(database = "DatabaseName", server = "ServerName\\Instance", file_path = "C:\\Temp\\Sdr\\ExcelFile.xlsx", archive_directory_path = "C:\\Temp\\Sdr\\ArchiveFolder\\")`

### Metadata

Return metadata for all tables in a database (table name, number of rows, size (in KB), and last schema update):

`adm_metadata_tables(database = "DatabaseName", server = "ServerName")`

Return metadata for all columns in a table (column name, data type, null count, distinct values count, minimum value, and maximum value):

`adm_metadata_columns(database = "DatabaseName", server = "ServerName", table = "TableName")`

### Versioning

To create a new version of a dataset, use `adm_create_version` specifying a `version_description` string, and a list of `version_tables` to include:

`adm_create_version(database = "DatabaseName", server = "server_name", version_description  = "A unique description of this version", version_table = c("Table1Name", "Table2Name"))`

To see a dataframe with details of all existing versions on a database:

`adm_list_versions(database = "DatabaseName", server = "serverName")`

To return a complete dataset from an existing version:

`adm_import_version(database = "DatabaseName", server = "ServerName", version_description = "The description given to this version")`

Future Development
---------
This package is currently under active development; feel free to add issues to this repo or contact the [ADM support team](mailto:adm.support@gov.scot) with any suggestions/problems.
