---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# sdrUpload

<!-- badges: start -->
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip) [![Travis-CI Build Status](https://api.travis-ci.org/thomascrines/admStructuredData.svg?branch=master)](https://travis-ci.org/thomascrines/admStucturedData)
<!-- badges: end -->

admStructuredData is an R package which allows SG analysts to upload data from Excel workbooks to the Analytical Data Management (ADM) structured data repository.

Data can be uploaded from a single worksheet, all worksheets in a workbook, or all workbooks in a folder. 

To use the package you will need to have access to a database in the ADM structured data repository; if you don't have this please contact the [ADM support team](mailto:adm.support@gov.scot).

## Installation

Install opendatascot from GitHub with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("thomascrines/admStructuredData")
```

If the above does not work, you can install from source:

1. Go to the sdrUpload [repository](https://github.com/thomascrines/admStructuredData) on GitHub
2. Click **Clone or download** then **Download ZIP**
3. Save the file locally and unzip
4. Install with install.packages():

```{r source-installation, eval = FALSE}
install.packages("C:\\DownloadDirectory\\admStructuredData-master\\admStructuredData-master", repos = NULL,
                 type="source", lib = "C:\\YourLibraryPath")
```

Prerequisites
---------
To use admStructuredData you will need to have access to a database in the ADM structured data repository; if you don't have this please contact the [ADM support team](mailto:adm.support@gov.scot). Most functions will require 'database' and 'server' arguments; the values for these will be provided by the ADM team when your database is set up.

Some functions allow processed files to be moved to an archive folder. To use this feature, the archive folder has to be made in advance.

Examples
---------
### Write data to database

Write an R dataframe to a database table:

`adm_dataframe_to_db(database = "DatabaseName", server = "ServerName", database_table_name = "example_database_table", dataframe = example_dataframe)`

Write data from an Excel worksheet to a database table:

`adm_worksheet_to_db(database = "DatabaseName", server = "ServerName", file_path = "C:\\Temp\\Sdr\\ExcelFile.xlsx", worksheet = "WorksheetName")`

Write data all Excel sheets contained in an Excel workbook to database tables:

`adm_workbook_to_db(database = "DatabaseName", server = "ServerName\\Instance", file_path = "C:\\Temp\\Sdr\\ExcelFile.xlsx")`

Write data from all Excel sheets contained in all Excel files contained in a source directory to database tables:

`adm_folder_to_db(database = "DatabaseName", server = "ServerName\\Instance", file_path = "C:\\Temp\\Sdr\\ExcelFile.xlsx")`

### Other database functions

Delete a table from a database:

`adm_delete_table(database = "DatabaseName", server = "ServerName", table = "TableName")`

Return a list of tables in a database:

`adm_list_tables(database = "DatabaseName", server = "DatabaseServer")`

Return data from a database table:

`adm_read_table(database = "DatabaseName", server = "ServerName", table = "TableName")`

### Append / overwrite arguments

`adm_dataframe_to_db()`, `adm_worksheet_to_db()`, `adm_workbook_to_db()` and `adm_folder_to_db()` all accept `append` and `overwrite` arguments.
These are optional, but if neither is set an error will be thrown if the database table already exists.
Only one can be set to `TRUE` at a time or an error will be thrown.

Append:

`adm_dataframe_to_db(database = "DatabaseName", server = "ServerName", database_table_name = "example_database_table", dataframe = example_dataframe, append = TRUE)`

Overwrite:

`adm_dataframe_to_db(database = "DatabaseName", server = "ServerName", database_table_name = "example_database_table", dataframe = example_dataframe, overwrite = TRUE)`

### Archive argument

`adm_worksheet_to_db()` and `adm_workbook_to_db()` can also accept an archive argument, to move processed files from the original location to a specified directory:

`adm_folder_to_db(database = "DatabaseName", server = "ServerName\\Instance", file_path = "C:\\Temp\\Sdr\\ExcelFile.xlsx", archive = "C:\\Temp\\Sdr\\ArchiveFolder\\")`

Future Development
---------
This package is currently under active development; feel free to add issues to this repo or contact the [ADM support team](mailto:adm.support@gov.scot) with any suggestions/problems.
